Integration of national demographic-disturbance relationships and local data can improve caribou population viability projections and inform monitoring decisions

### 2025-10-02 analysis/paper/paper.Rmd
Update in progress: main text figures have been updated. Associated text has not been revised.

resultScenario = "S13"
resultSet ="cmult6ay1aSf1"
figSet = c("diffslow","distScns","EVSI","power","EVPIall","powerLam","probViability")

@@

Figure 6: The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$, none), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). When disturbance is high the probability of correct status assessment is high even with one year of monitoring. Where disturbance is low or moderate, increasing the number of years of monitoring increases accuracy, but there are diminishing returns on additional monitoring as the number of years increases. In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year; varying these parameters does not alter qualitative conclusions (Supplements D and E).

Figure 7: The expected value of sample information (EVSI) varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$, none), and monitoring effort. When disturbance is high EVSI is low, indicating that local monitoring does not change predicted status. Where disturbance is lower, EVSI increases with the number of years of monitoring and the number of collars, but there are diminishing returns on additional monitoring as the number of years increases. See Figure \@ref(fig:figStatusErrors) for additional scenario details.

Figure 8: When anthropogenic disturbance is low and expected to increase (low scenario in Figure \@ref(fig:figDisturbance)) the probability of a correct status assessment varies with the true growth rate of the population (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$, none), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). Increasing the number of years of monitoring helps identify populations with currently declining populations (top left), and to identify populations with higher than expected growth rates that fare better as disturbance increases (lower right). Increasing the number of collars more effectively increases near-term accuracy when interannual variation is low (top left). If the goal is long-term projection (bottom row) there is little advantage to increasing the number of collars. Increasing the accuracy of status assessment for populations with growth rates close to one is difficult (middle columns). In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year.

Figure 9: The posterior probability that the population is viable ($P(\psi|x,\tau)$ in eq \@ref(eq:EVSI)) predicts the probability of a correct status assessment (estimated from the 114,000 example cases in Figures 5 to 7 with a generalized additive model). The relationship between accuracy and the probability of viability depends on true composition survey bias (colours), but regardless of bias accuracy is high when the probability of viability is very near to 0 or 1. Cases in which there is no need for additional monitoring if one is willing to accept a 5% error rate fall above the dashed line.

>>>>

resultScenario = "S3"
resultSet ="cmult6ay1aSf1"
figSet = c("diffslow","distScns","EVSI","power","EVPIall","powerLam","probViability")

@@

Figure 6: The probability of a correct status assessment varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). When disturbance is high the probability of correct status assessment is high even with one year of monitoring. Where disturbance is low or moderate, increasing the number of years of monitoring increases accuracy, but there are diminishing returns on additional monitoring as the number of years increases. In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year; varying these parameters does not alter qualitative conclusions (Supplements D and E).

Figure 7: The expected value of sample information (EVSI) varies with the amount of anthropogenic disturbance (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$), and monitoring effort. When disturbance is high EVSI is low, indicating that local monitoring does not change predicted status. Where disturbance is lower, EVSI increases with the number of years of monitoring and the number of collars, but there are diminishing returns on additional monitoring as the number of years increases. See Figure \@ref(fig:figStatusErrors) for additional scenario details.

Figure 8: When anthropogenic disturbance is low and expected to increase (low scenario in Figure \@ref(fig:figDisturbance)) the probability of a correct status assessment varies with the true growth rate of the population (columns), the number of years projected (rows), the amount of interannual variation (high: $v_R=0.46,v_S=0.087$, low: $v_R=0.23,v_S=0.0435$), and monitoring effort (see Figure \@ref(fig:figDisturbance) for scenarios). Increasing the number of years of monitoring helps identify populations with currently declining populations (top left), and to identify populations with higher than expected growth rates that fare better as disturbance increases (lower right). Increasing the number of collars more effectively increases near-term accuracy when interannual variation is low (top left). If the goal is long-term projection (bottom row) there is little advantage to increasing the number of collars. Increasing the accuracy of status assessment for populations with growth rates close to one is difficult (middle columns). In these examples the composition cow multiplier $w$ is 6 and collars are renewed each year.

Figure 9: The posterior probability that the population is viable ($P(\psi|x,\tau)$ in eq \@ref(eq:EVSI)) predicts the probability of a correct status assessment (estimated from the 76,000 example cases in Figures 5 to 7 with a generalized additive model). The relationship between accuracy and the probability of viability depends on true composition survey bias (colours), but regardless of bias accuracy is high when the probability of viability is very near to 0 or 1. Cases in which there is no need for additional monitoring if one is willing to accept a 5% error rate fall above the dashed line.
